\documentclass{article} \usepackage{fullpage} \title{Reding \emph{The Gender
Agenda in Computer Ethics} by Alison Adam} \author{Rikard Hjort}
\begin{document}
\maketitle

% Quick summary Adam's chapter acts as an introduction to feminist theory for
the field of computer ethics, and argues that feminist theory, and feminist
ethics in particular, can advance computer ethics. Adam makes a critical
analysis of some influential quantative studies on ethics, decision making, and
gender publihed in business and management literature. In particular, she is
critical of the unquestioned positivism that allows authors to hide possible
biases, e.g., relating to gender, behind a veil of objectivity. A pure
statistical analysis, Adam argues, hides how power relations could reasonably
influence the results.\footnote{p. 596}

The debate over quantative vs. qualitative methods has been raging among social
scientists for decades, and the pros and con have been thoroughly scrutinized.
However, this reasoning and justification of methodology is mostly missing in
the present studies\footnote{p. 597}, and as a result, their conclusions come
under question. The main issue, according to Adam, is an ``undertheorizing of
gender and ethics''\footnote{pp. 601}, partly caused by the fragmentation of the
research space -- the studies in question seem unaware of each other, as well as
of much of feminist research. This leads to some less informed conclusions being
drawn, seemingly based on gender stereotypes, which may not be at all warranted
by the statistical findings, while still seeming obvious to the authors, and
apparantely also their reviewers.

Adam goes on to give some remarks on how a feminist view of gender and computing
can help reforming computing to attract more women, and argues that this is a
valid goal, especially as the exclusion of women, wether intended or not, means
they are not present for conversations that impact society to high
degree.\footnote{This is the only part of the chapter I want to critizise: Adam
shows what I perceive to be a common bias, often present in discussion on gender
in the STEM subjects, viz. that it is assumed that the lack of women in STEM (in
this case computing) is their loss. Adam writes that ``women are still absent
from employment in well-paid and interesting careers'' [p. 605], a framing which
hides the fact that other well-paid and interesting careers with more human
interaction in fields such as law, governance and medicine, where women are not
underrepresented in the population entering these fields. It is far from sure,
but still important to consider, that that there may be skills that are more
prevalent among women that offer them more opportunities when deciding on a
career than their male peers, and that STEM simply does not rank that high for
people with many options.}

The sixth section of the chapter\footnote{pp. 606-611} is an application of feminist theory to
concrete issues in computer ethics: cyberstalking and hacking. It is a challenge
of the egalitarian ethos of the hacker community (hacking taken to mean both the
act of making software, and the more illicit activities of breaking into
computer systems). The main piece of work Adam scrutinizes is a book on hacker
communities, and Adam gives a fairly standard feminist critique of the work --
presenting women first and foremost in their relation to men, as wives,
dehumanizes them and obscures their capabilities, and
that stating the ethos that everyone is welcome and judged by their skill in a
community does not automatically make it egalitarian and meritocratic.

Adam concludes by viewing the field of ``feminist computer ethics'' from the
reverse angle: how computer ethics infused with feminist critique can enrich
feminist ethics.\footnote{pp. 611-615} Thanks to Donna Haraway's seminal paper \textit{A Cyborg
Manifesto}, there is already a large body of ideas in the intersection of
theories of gender, feminism and technology, especially around technology as an
enabler of a post-gender world. A popular term for the feminism at this
intersection is \textit{cyberfeminism}, and is largely centered around studying
female hackers as examples of women controlling technology, and thereby changing
power dynamics.

\paragraph{}The reason that I chose this chapter is that I have a background in
Gender Studies, and skimming the chapter it seemed to have an interesting
approach. I was somewhat disappointed in the rather broad and
basic treatment of the items of critique -- the text seemed to be intended
foremost as an introductory text of feminist critique and computer
ethics to scholars of both fields. I think it was a clear and did a good job of
articulating important ideas clearly. However, I am in the
unfortunate position as a reader of being familiar with the methods being
introduced, and perhaps it did not give me as much value for my time as some
other chapter might have.

My main takeaway is Adam's description\footnote{pp. 591-592} of Carol Gilligan's
book \textit{In a Different Voice}, and her result that men and women, for
whatever reason, tend to use different means to arrive at ethical decision,
women to a lesser degree leaning on law and Kantian moral philosophy, while men
rarely consider relational aspects when making such decisions. Apparently,
Gilligan spent spent some time refuting the work of Lawrence Kohlberg which was
published the year before, and his outline of moral development, with Kantian,
rule-based ethics being the highest form of ethics, which Gilligan argues is
actually very much a gendered hierarchy, which mostly justifies traditional
masculine ethics rather than a universal value system. I have myself been guilty
of falling into the trap of reasoning like Kohlberg, and it will be very healthy
to remember this critique when faced will ethical dilemmes, in whatever form, in
my practice as a computer scientist in (hopefully) diverse professional
contexts, and in my life in general. I will make sure to not dismiss relational
arguments, but take care to reasses my own stances, which tend to be more
rule-oriented.
\end{document}


TASK
http://www.cems.uwe.ac.uk/~pchatter/2011/pepi/The_Handbook_of_Information_and_Computer_Ethics.pdf

Choose a chapter that interests you, read it, and write a short report:

    Summarize the contents of the chapter Why do you find this chapter
interesting? Will this chapter be helpful in making ethical choices in real
situations?

I’m expecting a report of about one A4 page.

Chapters shortlist: * Online Anonymity * Responsibilities for Information on the
Internet * The Ethics of Cyberconflict * Regulation and Governance of the
Internet [24 pages] * Information Overload [23 pages] * Cencorship and Access to
Expression [17 pages] * The Gender Agenda in Computer Ethics [33 pages] * The
Digital Divide: A Perspective from the Future [19 pages]